---
title: "0119_Intro to TSA and R"
author: "Liwei Jiang"
date: "`r Sys.Date()`"
output: pdf_document
---

## Basic knowledge of R

length() #is same as ncol in R, not nrow

```{r cars}
summary(cars)
```

## Recap of Stats and TSA

Autocorrelation is important because if we have great autocorrelation, we don't need many covariates then.

For autocorrelation, we do many lags. Sometimes we will do 40 lags.

***Stationary process***: distribution of observations doesn't change along with time. Therefore, the mean, variance, and autocovariance functions remain constant. 

Question: If so, that means Yt will stay constant, right? 
Answer: I asked this question but Luana said no. We will cover this in Jan 24. Let's see. 

***White Noise Series***: independen, identically distributed (i.i.d) random variable. E(et) = 0, and autocorrelation & autocovariance = 0.

So, when doing modeling, check the distribution of residues. If mean and autocovariance are 0 or close to 0, that justifies the goodness of the model.

```{r}
lagged = zlag(dataset) #用来求lag的series的function
cor()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
