---
title: "0124_Intro to R Markdown"
author: "Liwei Jiang"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
###Introduction to TSA. 

How to push????

Stationary process:
***white Noise*** is an example of a stationary process

Autocorrelation Function (ACF): 
the correlation between Yt and Yt-h. 到底是就这两个数字，还是lag为h的两个series呢？

Partial Autocorrelation Fuction (PACF): 
How to get the correlation between Yt and Yt-h without the effect of intermediate terms 中间项 (Yt-1, Yt-2,...,Yt-h+1). So we can get a pure effect of Yt on Yt-h. Then we can make the estimate unbiased (think about a regression). 

***Some Notes***
ACF can show seasonality...
For h=0, ACF and PACF are always 1. 因为Yt = Yt
PACF is always smaller than ACF (effects are tuled out)

Auto-regressive: 
Regress on previous Y values to project future unknown Y values.

```{r}
##preparation work
library(lubridate)
library(ggplot2)
```
完成了准备工作，那就load数据啦
```{r}
#first check the current path to see how to edit file path
getwd()

#load dataset
## "."一个点就是简略代表了现在的在使用的path，之后直接再加上需要表明的就好了
## ".."两个点就是从现在getwd()的path中往上返回一层

df <- read.table(file="./Data/inflowtimeseries.txt",header=FALSE,skip=0)

df = df[, 1:5] #只取前五列

colnames(df) = c("Month", "Year", "HP1", "HP2", "HP3") #重命名前五列的列名

#简单的查看一下数据
head(df) ##前六列
str(df) #看看结构，都是什么类型的

```

## 如果整理dataset使其适合TSA，用lubridate()
```{r}

#接下来把前两列合成一列，sep是分隔符，合成了month-year的形式
my_date = paste(df[, 1], df[, 2], sep = "-")
#把合成的一列指定为日期类型，因为TSA要有日期啦
my_date = my(my_date)
#把合成的列加到原来的data frame里
df <- cbind(my_date,df[,3:5]) #这里把前两列删掉了
head(df)

#以下是一些怎么改变日期存储形式的function
format(my_date[1],format="%y-%m")
format(my_date[1],format="%Y-%m")
format(my_date[1],format="%Y-%b")
```

###接下来画图
```{r}
plot(df[,"HP1"], type = "l", colour = "red", ylab = "", xlab = "")
abline(h = mean(df[,"HP1"]), col = "blue")

#怎么overlay两个图
plot(df[,"HP1"], type = "l", col = "red", ylab = "", xlab = "")
lines(df[,"HP2"], type = "l", col = "red", ylab = "", xlab = "")

#ggplot
ggplot(data = df, aes(x = my_date, y = Inflow)) + 
  geom_line(aes(y = HP1, col = "HP1")) +
  geom_line(aes(y = HP2, col = "HP2")) +
  geom_line(aes(y = HP3, col = "HP3")) 

```

## 把dataframe转换为time series objects，这样好做TSA
```{r}
ts_df = ts(data = df[, 2:ncol(df)], start = c(1933, 1))

head(ts_df)
```

###求autocorrelation
```{r}
acf(ts_df[, 1], lag.max = 10) #最大的lag为10
pacf(ts_df[, 1])

```