---
title: '0126 Importing Data and Trend Modeling'
author: "Liwei Jiang"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r cars}
setwd("/Users/jiangliwei/Documents/Semester 4/Time Series Anaysis for Energy Data/Forked Repository/Forked TSA Spring 2023")
library(forecast)
library(Kendall)
library(tseries)
library(lubridate)
library(ggplot2)
library(readxl)
```

## Importing data

```{r pressure, echo=FALSE}
df <- read_xlsx(path = "./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx", col_names = TRUE, skip = 10, sheet = 1)
```

## Sources of variations
1. Calendar: seasons, holidays, and weekends
  Like seasonality: usage of electronic devices is constantly higher during weekends than weekdays
  ***R can automatically identify the length of seasonality once we determine the frequency, like 365days.***

2. Natural Calamities: earthquake, epidemic, flood, drought
***They should be incorporated in white noise, since they are random and we hope the mean of them is 0.***

## Trend Component: Linear or non-linear
1. Long-term tendency: increase or decrease
***We should always try to use linear trend first though it doesn't seem like linear will fit. because if linear can produce similar results as non-linear. Why bother ourselves with more complex methods?***

  ## Model linear trend:
  use Ordinary Least Square (OLS) method. Then we need to remove the trend, by use the ovservation to deduct fitted values. 
  
  ##Moving Average for Non-Linear Trend Estimation (Only usable for short term forecasting)
  Smooth out fluctuation in data to show a pattern or trend more clearly. 
  I have questions about determining order.
```{r}
##function to use for linear regression
df %>% lm(formula = Y ~ X, ...)
```

