---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2023"
subtitle: "Assignment 4 - Due date 02/17/23"
author: "Tony Jiang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: inline
---
  
## Directions
  
You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. And to do so you will need to fork our repository and link it to your RStudio.  

Once you have the file open on your local machine the first thing you will do is rename the file such that it includes your first and last name (e.g., "LuanaLima_TSA_A04_Sp23.Rmd"). Then change "Student Name" on line 4 with your name. 

Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Submit this pdf using Sakai. 

R packages needed for this assignment: "xlsx" or "readxl", "ggplot2", "forecast","tseries", and "Kendall". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here

library(readxl)
library(ggplot2)
library(forecast)
library(tseries)
library(Kendall)
library(tidyverse)
```

## Questions 

Consider the same data you used for A3 from the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx". The data comes from the US Energy Information and Administration and corresponds to the December 2022 Monthly Energy Review. For this assignment you will work only with the column "Total Renewable Energy Production". 


```{r}
#Importing data set - using xlsx package

renewables <- read_xlsx("./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx", 
                        skip = 10)

read_col_names <- read_xlsx(path = "./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx", 
                            col_names = FALSE, skip = 10, n_max = 1)

colnames(renewables) <- read_col_names

renewables <- renewables[-1,] #remove notes

renewables_sub <- renewables$`Total Renewable Energy Production`

renewables_sub = as.numeric(renewables_sub) 
  #covert the character type to numeric type

```


## Stochastic Trend and Stationarity Tests

### Q1

Difference the "Total Renewable Energy Production" series using function diff(). Function diff() is from package base and take three main arguments:
* *x* vector containing values to be differenced;
* *lag* integer indicating with lag to use;
* *differences* integer indicating how many times series should be differenced.

Try differencing at lag 1 only once, i.e., make `lag=1` and `differences=1`. Plot the differenced series Do the series still seem to have trend?

***Answer***: The plot does not seem to have a trend. The trend seems to fluctuate around 0.
```{r}

diff(renewables_sub, lag = 1, differences = 1) %>% plot(type = "l")


diff(renewables_sub, lag = 1, differences = 1) %>% summary() %>% print()
  #this is used to see the property of the differenced series.

```

### Q2

Now let's compare the differenced series with the detrended series you calculated on A3. In other words, for the "Total Renewable Energy Production" compare the differenced series from Q1 with the series you detrended in A3 using linear regression. (Hint: Just copy and paste part of your code for A3)

Copy and paste part of your code for A3 where you compute regression for Total Energy Production and the detrended Total Energy Production

```{r}

#ts_renewables_sub <- ts(renewables_sub, start = c(1973, 1), frequency = 12)

t = c(1:nrow(renewables))

# linear regression and store all coefficients
lm(renewables_sub ~ t) %>% coefficients() -> cfs

# detrend the series
detrend_renewables = renewables_sub - (cfs[[2]]*t + cfs[[1]])

# difference the series
differenced_renewables <- diff(renewables_sub, lag = 1, differences = 1)

```

### Q3
Create a data frame with 4 columns: month, original series, detrended by Regression Series and differenced series. Make sure you properly name all columns. Also note that the differenced series will have only 596 rows because you loose the first observation when differencing. Therefore, you need to remove the first observations for the original series and the detrended by regression series to build the new data frame.

```{r}
#Data frame - remember to not include January 1973

all_renewables <- data.frame(month = renewables[2:597,1], 
                             original = renewables_sub[2:597],
                             detrended = detrend_renewables[2:597],
                             differenced = differenced_renewables)
```

### Q4
Using ggplot() create a line plot that shows the three series together. Make sure you add a legend to the plot.

```{r}
# Use ggplot
# attached the data frame so don't need to type data frame name every time
attach(all_renewables)

# I convert the data frame into a long data frame so I can ggplot them
# easily by using group and colour
gather(all_renewables, key = "series", value = "values", 
       original, detrended, differenced) %>%
  ggplot(aes(x = Month, y = values, group = series, colour = series)) + 
  geom_line() +
  ylab("Renewable Energy Production") +
  xlab("Time") +
  ggtitle("Plots of All Three Series") +
  theme(plot.title = element_text(hjust = 0.5))


# codes below are the original code I used for this question. I leave it here
# to remind me of how ggplot runs

# ggplot(data = all_renewables, aes(x=Month)) +
#   geom_line(aes(y=original, color = "Original")) +
#   geom_line(aes(y=detrended, color = "Detrended")) +
#   geom_line(aes(y=differenced, color = "Differenced")) +
#   scale_color_manual(name = "Series",
#                      values = c("Original" = "black", 
#                                 "Detrended" = "red", 
#                                 "Differenced" = "orange"),
#                      labels = c("Detrended",
#                                 "Differenced",
#                                 "Oringinal")
#                      )

```


### Q5
Plot the ACF for the three series and compare the plots. Add the argument `ylim=c(-0.5,1)` to the Acf() function to make sure all three y axis have the same limits. Which method do you think was more efficient in eliminating the trend? The linear regression or differencing?

***Answer***: Differencing is apprently more efficient in eliminating the trend. Based on the acf plots, acf of differenced series is now randomly distributed between negative and positive and less statistically significant, while acf of detrended series is still statistically significantly positive, which still suggests strong non-stationarity and a trend over time.

```{r}

# Have them as ts objects, starting from 1973.2 because the first ob is omitted
ts_all_renewables <- ts(all_renewables[,2:4], 
                        start = c(1973, 2), frequency = 12)

#Compare ACFs

# names to use for plot titles
name_series = c("Original Series",
                "Detrended Series",
                "Differenced Series"
                )
# set the graph area to have three graphs displayed together
par(mfrow=c(1,3))

# plot their ACF plots
for (i in c(1:3)){
  
  acf(ts_all_renewables[,i], 
      main = paste("ACF of \n", name_series[i]),
      ylim = c(-1, 1)
      )
}

```



### Q6
Compute the Seasonal Mann-Kendall and ADF Test for the original "Total Renewable Energy Production" series. Ask R to print the results. Interpret the results for both test. Whats the conclusion from the Seasonal Mann Kendall test? What's the conclusion for the ADF test? Do they match what you observed in Q2? Recall that having a unit root means the series has a stochastic trend. And when a series has stochastic trend we need to use a different procedure to remove the trend.

***Answer***: 
The result of seasonal Kendall test is statistically significant, based on which we will reject the null hypothesis and say this series follows a trend. 
The result of Augmented Dick-Fuller test is statistically insignificant, based on which we fail to reject the null hypothesis (The series contains a unit root (non-stationarity).)

```{r}
#
sapply(renewables[, 5], as.numeric) %>% 
  ts(start = c(1973, 1), frequency = 12) -> ts_renewables_sub

# Seasonal Mann Kendall test and print the result
SeasonalMannKendall(ts_renewables_sub) %>% summary() %>% print()

# Augmented Dick-Fuller test and print the result
adf.test(ts_renewables_sub) %>% print()

```

### Q7

Aggregate the original "Total Renewable Energy Production" series by year. You can use the same procedure we used in class. Store series in a matrix where rows represent months and columns represent years. And then take the columns mean using function colMeans(). Recall the goal is the remove the seasonal variation from the series to check for trend. 

```{r}
#year 2022 is omitted by this method since 2022 doesn't have all 12 obs 
# to fill up the matrix
sapply(renewables[1:588, 5], as.numeric) %>% 
  matrix(nrow=12, byrow = FALSE, 
         dimnames = list(c(1:12), c(1973:2021))) -> years_renewables

# store column means to a vector
year_means <- colMeans(years_renewables)

```

### Q8

Apply the Mann Kendal, Spearman correlation rank test and ADF. Are the results from the test in agreement with the test results for the non-aggregated series, i.e., results for Q6?

**Answer***: Kendall test is statistically significant, based on which we will reject the null hypothesis and say yearly mean series follows a trend. 
Spearman correlation rank test is statistically significantly positive, which means the yearly means has a monotonic increasing trend over time. 
Augmented Dick-Fuller test is statistically insignificant, based on which we fail to reject the null hypothesis (The series contains a unit root.)

These results align with our results for the non-aggregated series. 

```{r}
# have a vector of time to conduct tests
years <- c(1:49)

# Kendall test
Kendall(x = years, y = year_means)

# Spearman correlation rank test
cor.test(x = years, y = year_means)

# Augmented Dick-Fuller test
adf.test(year_means)
```

