---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2023"
subtitle: "Assignment 2 - Due date 02/03/23"
author: "Tony Jiang"
output: pdf_document
geometry: margin=2.54cm
---

## R packages

R packages needed for this assignment:"forecast","tseries", and "dplyr". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here
library(forecast)
library(tseries)
library(dplyr)
```


## Data set information  

Consider the data provided in the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx" on our **Data** folder. The data comes from the US Energy Information and Administration and corresponds to the December 2022 Monthly Energy Review. The spreadsheet is ready to be used. You will also find a $.csv$ version of the data "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source-Edit.csv". You may use the function $read.table()$ to import the $.csv$ data in R. Or refer to the file "M2_ImportingData_CSV_XLSX.Rmd" in our Lessons folder for functions that are better suited for importing the $.xlsx$.

```{r}
#Importing data set and clean the data sets

df1 <- read_xlsx("./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx", col_names = TRUE, skip = 10)

df2 <- read.table(file="./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source-Edit.csv", header=TRUE,dec = ".",sep=",",stringsAsFactors = TRUE) 

df1 = df1[-1,]
```


## Question 1
You will work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series only. Use the command head() to verify your data.

```{r}
col_names = c("Month", "Total Biomass Energy Production", "Total Renewable Energy Production", "Hydroelectric Power Consumption")
df1_red = subset(df1, select = col_names)
head(df1_red)
```


## Question 2
Transform your data frame in a time series object and specify the starting point and frequency of the time series using the function ts().

```{r}
df1_red$Month = ymd(df1$Month)
ts_df1 = ts(df1_red, )
head(ts_df1)
```


## Question 3
Compute mean and standard deviation for these three series.

```{r}
a = c(colnames(ts_df1))
mean_list = data.frame(c("Mean of Biomass Energy Production", 
                   "Mean of Renewable Energy Production", 
                   "Mean of Hydroelectric Power Consumption"), 
                 c(mean(ts_df1[, "Total Biomass Energy Production"]),
                   mean(ts_df1[, "Total Renewable Energy Production"]),
                   mean(ts_df1[, "Hydroelectric Power Consumption"])
                   )
                 )
mean_list

std_list = data.frame(c("Std of Biomass Energy Production", 
                   "Std of Renewable Energy Production", 
                   "Std of Hydroelectric Power Consumption"), 
                 c(sd(ts_df1[, "Total Biomass Energy Production"]),
                   sd(ts_df1[, "Total Renewable Energy Production"]),
                   sd(ts_df1[, "Hydroelectric Power Consumption"])
                   )
                 )
std_list
```


## Question 4
Display and interpret the time series plot for each of these variables. Try to make your plot as informative as possible by writing titles, labels, etc. For each plot add a horizontal line at the mean of each series in a different color.

***Answers:*** 
1.The time series plot of biomass energy production generally shows an upward trend over time. But in the middle, the trend is stagnant. This may be caused by the change of regulation environment.
2. The time series plot of renewable energy production also shows an upward trend over time. Meanwhile, it shows seasonality within each year. 
3. The time series plot of hydroelectric power consumption shows great seasonality. It is hard to say whether it grows or decreases over time. But I suspect it decreases a little bit because, in recent years, the consumption is generally below the mean value. 

```{r}

for (i in c(2:4)){ 
plot(ts_df1[ , a[i]], type = "l",
     xlab = "Time",
     ylab = a[i])
abline(h=mean_list[[i-1,2]], col = "red")
title(main = paste("Time Series Plot of \n",a[i]), line = 0.2)
}

```


## Question 5
Compute the correlation between these three series. Are they significantly correlated? Explain your answer.

```{r}

```


## Question 6
Compute the autocorrelation function from lag 1 up to lag 40 for these three variables. What can you say about these plots? Do the three of them have the same behavior?

```{r}

```


## Question 7
Compute the partial autocorrelation function from lag 1 to lag 40 for these three variables. How these plots differ from the ones in Q6?

```{r}

```


