---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2023"
subtitle: "Assignment 3 - Due date 02/10/23"
author: "Tony Jiang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: inline
---
  
## Directions


Please keep this R code chunk options for the report. It is easier for us to grade when we can see code and output together. And the tidy.opts will make sure that line breaks on your code chunks are automatically added for better visualization. 

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

When you have completed the assignment, **Knit** the text and code into a single PDF file. Submit this pdf using Sakai. 

## Questions 

Consider the same data you used for A2 from the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx". The data comes from the US Energy Information and Administration and corresponds to the December 2022 **Monthly** Energy Review. Once again you will work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series only. 

R packages needed for this assignment:"forecast","tseries", and "Kendall". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here
library(forecast)
library(tseries)
library(Kendall)
library(xlsx)
```

###Load the dataset

```{r}
renewables <- read.xlsx("./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx", 
                        header = TRUE, startRow = 11, sheetIndex=1)

read_col_names <- read.xlsx(file="./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",header=FALSE,startRow = 11,endRow=11,sheetIndex=1)

colnames(renewables) <- read_col_names

renewables <- renewables[-1,] 

col_names = c("Total Biomass Energy Production", 
              "Total Renewable Energy Production", 
              "Hydroelectric Power Consumption")

renewables_sub <- renewables[, 4:6]

head(renewables_sub)
```

###Coerce the dataset as time series data
```{r}
renewables_sub[, c(1:3)] = sapply(renewables_sub[, c(1:3)], as.numeric) #convert columns to numeric. Otherwise numbers will be altered by ts().

ts_renewables <- ts(renewables_sub, start = c(1973, 1), frequency = 12)

head(ts_renewables)
```


##Trend Component

### Q1
Create a plot window that has one row and three columns. And then for each object on your data frame, fill the plot window with time series plot, ACF and PACF. You may use the some code form A2, but I want all three plots on the same window this time. (Hint: use par() function)

```{r}

for (i in 1:3){
  par(mfrow = c(1,3))
  plot(ts_renewables[,i], main = paste("Time Series Plot of \n", col_names[i]), 
       xlab = "Years", ylab = col_names[i])
  acf(ts_renewables[,i], lag.max = 40, main = paste("Autocorrelation of \n", col_names[i]), ylim = c(-1, 1))
  pacf(ts_renewables[,i], lag.max = 40, main = paste("Autocorrelation of \n", col_names[i]), ylim = c(-1, 1))
}

```


### Q2
From the plot in Q1, do the series Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption appear to have a trend? If yes, what kind of trend?

***Answers to Q2***
Yes. Total Biomass Energy Production & Total Renewable Energy Production display significant upward trends. Hydroelectric Power Consumption appears to have a downward trend but not significant. 
As for seasonality, Hydroelectric Power Consumption has a strong seasonality. The other two series don't show obvious seasonality. But the ACF and PACF of these three series are statistically significant which indicates their non-stationarity. (Their distribution properties change over time)


### Q3
Use the *lm()* function to fit a linear trend to the three time series. Ask R to print the summary of the regression. Interpret the regression output, i.e., slope and intercept. Save the regression coefficients for further analysis.

```{r}
t <- c(1:nrow(ts_renewables))

cfs <- list()

for (i in 1:3){
  lm(ts_renewables[, i] ~ t) %>% summary() %>% print()
  
  lm(ts_renewables[, i] ~ t) %>% coefficients() -> cf_raw
  cfs <- append(cfs, cf_raw)
}
```
 
### Q4
Use the regression coefficients from Q3 to detrend the series. Plot the detrended series and compare with the plots from Q1. What happened? Did anything change?

```{r}
remove(detrend_renewables)
de_names = c(paste("Detrended", col_names[1]),
             paste("Detrended", col_names[2]),
             paste("Detrended", col_names[3])
             )

detrend_renewables = data.frame(c(1:nrow(ts_renewables)),
                                c(1:nrow(ts_renewables)),
                                c(1:nrow(ts_renewables))
                                )

colnames(detrend_renewables) = de_names

for (i in 1:3){
  
  detrend_renewables[,i] = ts_renewables[,i] - (cfs[[2*i]]*ts_renewables[,i] + cfs[[2*i-1]])
  
}

ts_detrend_renewables = ts(detrend_renewables, start = c(1973, 1), frequency = 12)

head(ts_detrend_renewables)
```


### Q5
Plot ACF and PACF for the detrended series and compare with the plots from Q1. Did the plots change? How?

```{r}

for (i in 1:3){
  par(mfrow = c(1,3))
  plot(ts_detrend_renewables[,i], main = paste("Time Series Plot of \n Detrended", col_names[i]), 
       xlab = "Years", ylab = col_names[i])
  
  acf(ts_detrend_renewables[,i], lag.max = 40, 
      main = paste("Autocorrelation of \n Detrended", col_names[i]), 
      ylim = c(-1, 1)
      )
  
  pacf(ts_detrend_renewables[,i], lag.max = 40, 
       main = paste("Autocorrelation of \n Detrended", col_names[i]), 
       ylim = c(-1, 1)
       )
}

```


## Seasonal Component

Set aside the detrended series and consider the original series again from Q1 to answer Q6 to Q8.

### Q6
Do the series seem to have a seasonal trend? Which serie/series? Use function *lm()* to fit a seasonal means model (i.e. using the seasonal dummies) to this/these time series. Ask R to print the summary of the regression. Interpret the regression output. Save the regression coefficients for further analysis.

### Q7
Use the regression coefficients from Q6 to deseason the series. Plot the deseason series and compare with the plots from part Q1. Did anything change?

### Q8
Plot ACF and PACF for the deseason series and compare with the plots from Q1. Did the plots change? How?


